---
title: "GDD-Plots-2022"
output: html_document
date: '2022-05-11'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Web Scraping

### West Madison

#### Temperature

Using the Middleton Airport data (KC29) for temperature data

There is no rainfall data included in this, unfortunately

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(curl)
library(rjson)

curl_download("https://api.weather.gov/stations/KC29/observations", "KC29.json")

data <- fromJSON(file="KC29.json")

times <- data.frame(row=factor(),
                    time = character(),
                    temp=double(),
                    precipitation=double())

for (i in 1:length(data$features)) {
  if (length((data$features[[i]])$properties$temperature$value) != 0) {
    tempdf <- data.frame(i, (data$features[[i]])$properties$timestamp, (data$features[[i]])$properties$temperature$value)
    names(tempdf) <- c("row", "time", "temp")
    times <- rbind(times, tempdf)
  }
}

times <- times %>%
  separate(time, c("Day", "Time"), sep="T") %>%
  mutate(Time = substr(Time, 1,5)) %>%
  separate(Day, c("Year", "Month", "Day")) %>%
  mutate(tempF = temp * (9/5) + 32) %>%
  mutate(Date = paste(Month, Day, Year, sep="/")) %>%
  group_by(Date) %>%
  summarize(MaxTemp = max(tempF),
            MinTemp = min(tempF))
  

write.csv(times, "WMARS.csv")

```

#### Rainfall

Using the UW-Madison Charmany Farm (CHMW3, Vet School facility on Mineral Point Road) for precipitation data

see here: https://mesonet.agron.iastate.edu/request/coop/obs-fe.phtml?network=WI_COOP

```{r}
curl_download("https://mesonet.agron.iastate.edu/request/coop/obs-dl.php?network=WI_COOP&station%5B%5D=CHMW3&year1=2023&month1=4&day1=10&year2=2023&month2=12&day2=1&what=view&delim=comma", "CHMW3.csv")
```

The values are in inches. Same for the Michigan data


### Hancock

Michigan has changed its website in 2022, so I added `legacy` to the beginnings of the URLs

Year is hard coded in here!

```{r}
library(rvest)

Date <- Sys.Date()

data <- read_html(paste("https://legacy.enviroweather.msu.edu/run.php?stn=hck&mod=w_sum&da1=28&mo1=4&da2=", (as.numeric(substr(Date, 9, 10))-1), "&mo2=", substr(Date, 6, 7), "&yr=2023&mc=604&ds=cd", sep="")) %>% 
  html_elements("td") %>%
  html_text2()

matrix <- t(replace(matrix(NA, 17, ceiling(length(data) / 17)), seq_along(data), data))
data <- as.data.frame(matrix)

write.table(data, "Hancock.csv", sep=",", col.names=FALSE, row.names=FALSE)

```

### Arlington

```{r}
library(rvest)

Date <- Sys.Date()

data <- read_html(paste("https://legacy.enviroweather.msu.edu/run.php?stn=alt&mod=w_sum&da1=10&mo1=5&da2=", (as.numeric(substr(Date, 9, 10))-1), "&mo2=", substr(Date, 6, 7), "&yr=2023&mc=604&ds=cd", sep="")) %>% 
  html_elements("td") %>%
  html_text2()

matrix <- t(replace(matrix(NA, 17, ceiling(length(data) / 17)), seq_along(data), data))
data <- as.data.frame(matrix)

write.table(data, "Arlington.csv", sep=",", col.names=FALSE, row.names=FALSE)

```

### Marshfield

```{r}
library(rvest)

Date <- Sys.Date()

data <- read_html(paste("https://legacy.enviroweather.msu.edu/run.php?stn=mar&mod=w_sum&da1=10&mo1=5&da2=", (as.numeric(substr(Date, 9, 10))-1), "&mo2=", substr(Date, 6, 7), "&yr=2023&mc=604&ds=cd", sep="")) %>% 
  html_elements("td") %>%
  html_text2()

matrix <- t(replace(matrix(NA, 17, ceiling(length(data) / 17)), seq_along(data), data))
data <- as.data.frame(matrix)

write.table(data, "Marshfield.csv", sep=",", col.names=FALSE, row.names=FALSE)

```

## Updating the sheets

If the authentication doesn't work, usually, `gs4_auth()` needs to be run.

### The GDD Equation

```{r}
calc.daily.GDD <- function (max, min) {
  min <- ifelse(min < 50, 50, min)
  max <- ifelse(max > 86, 86, max)
  mean <- (min + max) / 2
  ifelse(mean < 50, 0, mean - 50)
}
```



### Hancock

```{r}
library(lubridate)
library(googlesheets4)
'%!in%' <- function(x,y)!('%in%'(x,y))

# Initialize the spreadsheet for the season manually if it hasn't been already

if (!file.exists("Hancock_existing.csv")) {
  WMARS_existing <- read_csv(file = "Date, Max, Min, GDDDaily
                             05/16/2022, 72.1, 48.9, NA")
  write.csv(WMARS_existing, "Hancock_existing.csv")
}

Existing <- read.csv("Hancock_existing.csv") %>%
  mutate(JulianDay = yday(as.Date(Date, "%m/%d/%y")))

New <- read.csv("Hancock.csv") %>%
  mutate(Date = paste(Date, "/2022", sep="")) %>%
  mutate(JulianDay = yday(as.Date(Date, "%m/%d/%y"))) %>%
  filter(JulianDay %!in% Existing$JulianDay) %>%
  filter(JulianDay >= 137)

New <- merge(Existing, New, sort=FALSE, all=TRUE) %>%
  mutate(GDDDaily = calc.daily.GDD(max = Max, min = Min)) %>%
  filter(JulianDay >= 137) %>%
#  filter(JulianDay < yday(Sys.Date())) %>%
  arrange(JulianDay) %>%
  select(Date, Max, Min, GDDDaily)

for (row in 1:nrow(New)) {
  if (row == 1) {
    New[row, "K25-K28"] <- New[1, "GDDDaily"]
  }
  else {
    New[row, "K25-K28"] <- New[(row-1), "K25-K28"] + New[row, "GDDDaily"]
  }
}

for (row in 2:nrow(New)) {
  if (row == 2) {
    New[row, "K29-K32"] <- New[2, "GDDDaily"]
  }
  else {
    New[row, "K29-K32"] <- New[(row-1), "K29-K32"] + New[row, "GDDDaily"]
  }
}

for (row in 3:nrow(New)) {
  if (row == 3) {
    New[row, "S12-S16"] <- New[3, "GDDDaily"]
  }
  else {
    New[row, "S12-S16"] <- New[(row-1), "S12-S16"] + New[row, "GDDDaily"]
  }
}




sheet_write(New, "1_QmalVr6MfYJ7xmk6pPng4dfTFPoGuEu0qGVzhs0ryE", "HAN_automatic")

write.csv(New, "Hancock_existing.csv")

```



## Plots

### Grain Trials

```{r}
WMARSExisting <- read.csv("WMARS_existing.csv") %>%
  select(Date, M800) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

HANExisting <- read.csv("Hancock_existing.csv", check.names = FALSE) %>%
  select(Date, `K25-K28`) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

ARLExisting <- read.csv("Arlington_existing.csv", check.names = FALSE) %>%
  select(Date, `1E`, `712`) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

Merged <- merge(WMARSExisting, HANExisting, by = "Date", all=TRUE) %>%
  merge(ARLExisting, by = "Date", all = TRUE) %>%
  gather(M800, `K25-K28`, `1E`, `712`, key = "Field", value = "GDDs") %>%
  mutate(Field = case_when(
    Field == "M800" ~ "West Madison (M800)",
    Field == "K25-K28" ~ "Hancock (K25 - K28)",
    Field == "1E" ~ "Arlington (1E)",
    Field == "712" ~ "Arlington (712)"
  ))

Merged$Field <- factor(Merged$Field, levels = c("West Madison (M800)", "Hancock (K25 - K28)", "Arlington (1E)", "Arlington (712)"))

ggplot(Merged, aes(Date, GDDs, group=Field, color=Field)) +
  geom_line() +
  theme_classic() +
  ylab("Cumulative GDDs (ºF)") +
  ggtitle("Grain Trials") +
  geom_hline(yintercept = 1200, linetype = 'dotted', col = 'blue') +
  annotate("text", x = as.Date("06/20/2022", "%m/%d/%y"), y = 1350, label = "~ start of flowering") +
  geom_hline(yintercept = 120, linetype = 'dotted', col = 'blue') +
  scale_color_manual(values = c("#F8766D", "#00BA38", "#619CFF", "black")) + 
  geom_hline(yintercept = 2700, linetype = 'dotted', col = 'black') +
  annotate("text", x = as.Date("06/30/2022", "%m/%d/%y"), y = 2550, label = "~ physiological maturity")

ggsave("TrialGDDs.png", dpi=200, width = 6, height = 4, units = "in")
```


### Silage

```{r}

WMARSExisting <- read.csv("WMARS_existing.csv") %>%
  select(Date, M901) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

ARLExisting <- read.csv("Arlington_existing.csv", check.names = FALSE) %>%
  select(Date, `26N`) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

Merged <- merge(WMARSExisting, ARLExisting, by = "Date", all = TRUE) %>%
  gather(M901, `26N`, key = "Field", value = "GDDs") %>%
  mutate(Field = case_when(
    Field == "M901" ~ "West Madison (M901)",
    Field == "26N" ~ "Arlington (26N)"
  ))

Merged$Field <- factor(Merged$Field, levels = c("West Madison (M901)", "Arlington (26N)"))

ggplot(Merged, aes(Date, GDDs, group=Field, color=Field)) +
  geom_line() +
  theme_classic() +
  ylab("Cumulative GDDs (ºF)") +
  ggtitle("Silage Trials") +
  geom_hline(yintercept = 2195, linetype = 'dotted', col = 'blue') +
  annotate("text", x = as.Date("06/30/2022", "%m/%d/%y"), y = 2050, label = "~ 35% dry matter silage") +
  scale_color_manual(values = c("#00BA38", "#619CFF", "black")) #+
  #scale_x_date(limits = c(as.Date("05/9/2022", "%m/%d/%y"), as.Date("05/25/2021", "%m/%d/%y")))



ggsave("Silage.png", dpi=200, width = 6, height = 4, units = "in")

```

### Nursery

```{r}
WMARSExisting <- read.csv("WMARS_existing.csv") %>%
  select(Date, H102, B504, B505, P200) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y")) %>%
  gather(H102, B504, B505, P200, key = "Field", value = "GDDs")

WMARSExisting$Field <- factor(WMARSExisting$Field, levels = c("H102", "B504", "B505", "P200"))

ggplot(WMARSExisting, aes(Date, GDDs, group=Field, color=Field)) +
  geom_line() + 
  theme_classic() +
  ylab("Cumulative GDDs (ºF)") +
  ggtitle("Nurseries and Observations") +
  geom_hline(yintercept = 1200, linetype = 'dotted', col = 'blue') +
  annotate("text", x = as.Date("06/10/2022", "%m/%d/%y"), y = 1100, label = "~ start of flowering") +
  geom_hline(yintercept = 120, linetype = 'dotted', col = 'blue') +
  geom_hline(yintercept = 2700, linetype = 'dotted', col = 'black') +
  annotate("text", x = as.Date("06/10/2022", "%m/%d/%y"), y = 2550, label = "~ physiological maturity")

ggsave("NurseriesGDDs.png", dpi=500, width = 6, height = 4, units = "in")
```

### Flowering Progress

```
M800 <- read.csv("~/gem-fact-field/Data2022/GEMN FACT DH X LH244 2022 WM Phenotypes.csv")

M800Flowering <- data.frame(GDDs = unique(M800$AnthesisGDD),
                            ProportionFlowered = NA,
                            Field = "M800")

for (row in 1:nrow(M800Flowering)) {
  ThisValue <- M800Flowering[row, "GDDs"]
  tempdf <- M800 %>%
    filter(!is.na(AnthesisGDD)) %>%
    filter(AnthesisGDD <= ThisValue)
  M800Flowering[row, "ProportionFlowered"] <- nrow(tempdf) / nrow(M800)

}

PHP02 <- read.csv("~/gem-fact-field/Data2022/GEMS FACT DH X PHP02 2022 ARL Phenotypes.csv")

MBS5411 <- read.csv("~/gem-fact-field/Data2022/GEMS FACT DH X MBS5411 2022 ARL Phenotypes.csv")

LH244 <- read.csv("~/gem-fact-field/Data2022/GEMN FACT DH X LH244 2022 ARL Phenotypes.csv")

ARL712 <- rbind(PHP02, MBS5411, LH244)

ARL712Flowering <- data.frame(GDDs = unique(ARL712$AnthesisGDD),
                            ProportionFlowered = NA,
                            Field = "712") %>%
  filter(!is.na(GDDs))

for (row in 1:nrow(ARL712Flowering)) {
  ThisValue <- ARL712Flowering[row, "GDDs"]
  tempdf <- ARL712 %>%
    filter(!is.na(AnthesisGDD)) %>%
    filter(AnthesisGDD <= ThisValue)
  ARL712Flowering[row, "ProportionFlowered"] <- nrow(tempdf) / nrow(ARL712)

}

MergedFlowering <- rbind(M800Flowering, ARL712Flowering)

ggplot(MergedFlowering, aes(GDDs, ProportionFlowered, color = Field)) +
  geom_line() +
  xlab("Cumulative GDDs (ºF)") +
  ylab("Proportion of plots at anthesis") +
  ggtitle("Flowering Progress") +
  theme_classic()

ggsave("FloweringProgress.png", dpi=200, width = 6, height = 4, units = "in")
```


Here's the code to run this from the command line
```
Rscript -e "rmarkdown::render('GDD-Plots-2022.Rmd')"
```