---
title: "GDD-Plots-2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Web Scraping

### West Madison

#### Temperature

Using the Middleton Airport data (KC29) for temperature data

There is no rainfall data included in this, unfortunately

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(curl)
library(rjson)

curl_download("https://api.weather.gov/stations/KC29/observations", "KC29.json")

data <- fromJSON(file="KC29.json")

times <- data.frame(row=factor(),
                    time = character(),
                    temp=double(),
                    precipitation=double())

for (i in 1:length(data$features)) {
  if (length((data$features[[i]])$properties$temperature$value) != 0) {
    tempdf <- data.frame(i, (data$features[[i]])$properties$timestamp, (data$features[[i]])$properties$temperature$value)
    names(tempdf) <- c("row", "time", "temp")
    times <- rbind(times, tempdf)
  }
}

times <- times %>%
  separate(time, c("Day", "Time"), sep="T") %>%
  mutate(Time = substr(Time, 1,5)) %>%
  separate(Day, c("Year", "Month", "Day")) %>%
  mutate(tempF = temp * (9/5) + 32) %>%
  mutate(Date = paste(Month, Day, Year, sep="/")) %>%
  group_by(Date) %>%
  summarize(MaxTemp = max(tempF),
            MinTemp = min(tempF))
  

write.csv(times, "WMARS.csv")

```

#### Rainfall

Using the UW-Madison Charmany Farm (CHMW3, Vet School facility on Mineral Point Road) for precipitation data

see here: https://mesonet.agron.iastate.edu/request/coop/obs-fe.phtml?network=WI_COOP

```{r}
curl_download("https://mesonet.agron.iastate.edu/request/coop/obs-dl.php?network=WI_COOP&station%5B%5D=CHMW3&year1=2023&month1=4&day1=10&year2=2023&month2=12&day2=1&what=view&delim=comma", "CHMW3.csv")
```

The values are in inches. Same for the Michigan data


### Hancock

Michigan has changed its website in 2022, so I added `legacy` to the beginnings of the URLs

Year is hard coded in here!

```{r}
library(rvest)

Date <- Sys.Date()

data <- read_html(paste("https://legacy.enviroweather.msu.edu/run.php?stn=hck&mod=w_sum&da1=28&mo1=4&da2=", (as.numeric(substr(Date, 9, 10))-1), "&mo2=", substr(Date, 6, 7), "&yr=2023&mc=604&ds=cd", sep="")) %>% 
  html_elements("td") %>%
  html_text2()

matrix <- t(replace(matrix(NA, 17, ceiling(length(data) / 17)), seq_along(data), data))
data <- as.data.frame(matrix)

data[1,16] <- "Rain"

write.table(data, "Hancock.csv", sep=",", col.names=FALSE, row.names=FALSE)

```

### Arlington



### Marshfield

```{r}
library(rvest)

Date <- Sys.Date()

data <- read_html(paste("https://legacy.enviroweather.msu.edu/run.php?stn=mar&mod=w_sum&da1=10&mo1=5&da2=", (as.numeric(substr(Date, 9, 10))-1), "&mo2=", substr(Date, 6, 7), "&yr=2023&mc=604&ds=cd", sep="")) %>% 
  html_elements("td") %>%
  html_text2()

matrix <- t(replace(matrix(NA, 17, ceiling(length(data) / 17)), seq_along(data), data))
data <- as.data.frame(matrix)

data[1,16] <- "Rain"

write.table(data, "Marshfield.csv", sep=",", col.names=FALSE, row.names=FALSE)

```

## Updating the sheets

If the authentication doesn't work, usually, `gs4_auth()` needs to be run.

### The GDD Equation

```{r}
calc.daily.GDD <- function (max, min) {
  min <- ifelse(min < 50, 50, min)
  max <- ifelse(max > 86, 86, max)
  mean <- (min + max) / 2
  ifelse(mean < 50, 0, mean - 50)
}
```



### Hancock

```{r}
library(lubridate)
library(googlesheets4)
'%!in%' <- function(x,y)!('%in%'(x,y))

AddField <- function(data, Name, StartRow) {
  for (row in StartRow:nrow(data)) {
    if (row == StartRow) {
      data[row, Name] <- data[StartRow, "GDDDaily"]
   }
    else {
      data[row, Name] <- data[(row-1), Name] + data[row, "GDDDaily"]
    }
  }
  return(data)
}

CumulativeRain <- function(data) {
  for (row in 1:nrow(data)) {
    if(row == 1) {
      data[row, "Cumulative Rainfall (in)"] <- data[row, "Rain"]
    }
    else {
      data[row, "Cumulative Rainfall (in)"] <- data[(row-1), "Cumulative Rainfall (in)"] + data[row, "Rain"]
    }
  }
  return(data)
}

# Initialize the spreadsheet for the season manually if it hasn't been already

if (!file.exists("Hancock_existing.csv")) {
  existing <- read_csv(file = "Date, Max, Min, Rain, GDDDaily
                             4/25/2023, 49.2, 31.5, 0, NA")
  write.csv(existing, "Hancock_existing.csv")
}

Existing <- read.csv("Hancock_existing.csv") %>%
  mutate(JulianDay = yday(as.Date(Date, "%m/%d/%y")))

New <- read.csv("Hancock.csv") %>%
  mutate(Date = paste(Date, "/2023", sep="")) %>%
  mutate(JulianDay = yday(as.Date(Date, "%m/%d/%y"))) %>%
  filter(JulianDay %!in% Existing$JulianDay) %>%
  filter(JulianDay > 115)

New <- merge(Existing, New, sort=FALSE, all=TRUE) %>%
  mutate(GDDDaily = calc.daily.GDD(max = Max, min = Min)) %>%
  filter(JulianDay >= 115) %>%
#  filter(JulianDay < yday(Sys.Date())) %>%
  arrange(JulianDay) %>%
  select(Date, Max, Min, Rain, GDDDaily) %>%
  CumulativeRain() %>%
  AddField("S13-S16", 1) %>%
  AddField("S21-S24", 2) %>%
  AddField("S9-S12", 3) %>%
  AddField("S17-S20", 3)

sheet_write(New, "1_QmalVr6MfYJ7xmk6pPng4dfTFPoGuEu0qGVzhs0ryE", "HAN_automatic")

write.csv(New, "Hancock_existing.csv")

```


### WMARS

```{r}
library(lubridate)
library(googlesheets4)
'%!in%' <- function(x,y)!('%in%'(x,y))

# Initialize the spreadsheet for the season manually if it hasn't been already

if (!file.exists("WMARS_existing.csv")) {
  WMARS_existing <- read_csv(file = "Date, MaxTemp, MinTemp, Rain, GDDDaily
                             05/03/2023, 63.14, 33.62, 0, NA")
  write.csv(WMARS_existing, "WMARS_existing.csv")
}

Existing <- read.csv("WMARS_existing.csv") %>%
  mutate(JulianDay = yday(as.Date(Date, "%m/%d/%y")))

NewRain <- read.csv("CHMW3.csv") %>%
  mutate(JulianDay = yday(date) + 1) %>% #Julian days don't match WM for reasons I don't know
  select(JulianDay, Rain = precip)

New <- read.csv("WMARS.csv") %>%
  mutate(JulianDay = yday(as.Date(Date, "%m/%d/%y"))) %>%
  merge(NewRain, by = "JulianDay") %>%
  filter(JulianDay %!in% Existing$JulianDay) %>%
  mutate(Rain = as.numeric(Rain)) %>%
  mutate(Rain = case_when(
    is.na(Rain) ~ 0,
    Rain == "" ~ 0,
    TRUE ~ Rain
  )) # Sometimes there are missing values


New <- merge(Existing, New, sort=FALSE, all=TRUE) %>%
  mutate(GDDDaily = calc.daily.GDD(max = MaxTemp, min = MinTemp)) %>%
  filter(JulianDay > 123) %>%
  filter(JulianDay <= yday(Sys.Date())) %>% #This isn't quite right
  arrange(JulianDay) %>%
  select(Date, MaxTemp, MinTemp, Rain, GDDDaily) %>%
  CumulativeRain() %>%
  AddField("M1400", 1) %>%
  AddField("M1500", 2) %>%
  AddField("H703", 2) %>%
  AddField("B501", 20) %>%
  AddField("B502", 20) %>%
  AddField("H101", 20) %>%
  AddField("B601", 29) %>%
  AddField("H702", 30) %>%
  AddField("CERCA", 130)


sheet_write(New, "1_QmalVr6MfYJ7xmk6pPng4dfTFPoGuEu0qGVzhs0ryE", "WM_automatic")

write.csv(New, "WMARS_existing.csv")
```

### Arlington

```{r}
library(lubridate)
library(googlesheets4)
'%!in%' <- function(x,y)!('%in%'(x,y))


# Initialize the spreadsheet for the season manually if it hasn't been already

if (!file.exists("Arlington_existing.csv")) {
  existing <- read_csv(file = "Date, Max, Min, Rain, GDDDaily
                             5/11/2023, 81, 45.4, 0, NA")
  write.csv(existing, "Arlington_existing.csv")
}

Existing <- read.csv("Arlington_existing.csv") %>%
  mutate(JulianDay = yday(as.Date(Date, "%m/%d/%y")))

New <- read.csv("ARLVisualCrossing.csv") %>%
  mutate(Date = as.character(as.Date(datetime))) %>%
  separate(Date, c("Year", "Month", "Day")) %>%
  mutate(Date = paste(Month, Day, Year, sep="/")) %>%
  mutate(JulianDay = yday(as.Date(Date, "%m/%d/%y"))) %>%
  filter(JulianDay %!in% Existing$JulianDay) %>% # Didn't use to be necessary, but there is some stochasticity in the values the API gives back. Therefore, they might not match perfectly
  select(Date, Max = tempmax, Min = tempmin, Rain = precip, JulianDay)


New <- merge(Existing, New, sort=FALSE, all=TRUE) %>%
  mutate(GDDDaily = calc.daily.GDD(max = Max, min = Min)) %>%
  filter(JulianDay >= 130) %>%
  filter(JulianDay < yday(Sys.Date()) + 1) %>% # Still don't understand the +1
  arrange(JulianDay) %>%
  select(Date, Max, Min, Rain, GDDDaily) %>%
  CumulativeRain() %>%
  AddField("593", 1) %>%
  AddField("25N", 2) %>%
  AddField("28E", 2) %>%
  AddField("713", 7)

sheet_write(New, "1_QmalVr6MfYJ7xmk6pPng4dfTFPoGuEu0qGVzhs0ryE", "ARL_automatic")

write.csv(New, "Arlington_existing.csv")

```

### Marshifeld

```{r}
library(lubridate)
library(googlesheets4)
'%!in%' <- function(x,y)!('%in%'(x,y))


# Initialize the spreadsheet for the season manually if it hasn't been already

if (!file.exists("Marshfield_existing.csv")) {
  existing <- read_csv(file = "Date, Max, Min, Rain, GDDDaily
                             5/25/2023, 67.8, 39, 0, NA")
  write.csv(existing, "Marshfield_existing.csv")
}

Existing <- read.csv("Marshfield_existing.csv") %>%
  mutate(JulianDay = yday(as.Date(Date, "%m/%d/%y")))

New <- read.csv("Marshfield.csv") %>%
  mutate(Date = paste(Date, "/2023", sep="")) %>%
  mutate(JulianDay = yday(as.Date(Date, "%m/%d/%y"))) %>%
  filter(JulianDay %!in% Existing$JulianDay) %>%
  filter(JulianDay > 145)

New <- merge(Existing, New, sort=FALSE, all=TRUE) %>%
  mutate(GDDDaily = calc.daily.GDD(max = Max, min = Min)) %>%
  filter(JulianDay >= 145) %>%
#  filter(JulianDay < yday(Sys.Date())) %>%
  arrange(JulianDay) %>%
  select(Date, Max, Min, Rain, GDDDaily) %>%
  CumulativeRain() %>%
  AddField("Marshfield", 1)

sheet_write(New, "1_QmalVr6MfYJ7xmk6pPng4dfTFPoGuEu0qGVzhs0ryE", "MAR_automatic")

write.csv(New, "Marshfield_existing.csv")

```


## Plots

### Grain Trials

```{r}
HANExisting <- read.csv("Hancock_existing.csv", check.names = FALSE) %>%
  select(Date, `S13-S16`) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

WMARSExisting <- read.csv("WMARS_existing.csv", check.names = FALSE) %>%
  select(Date, M1400) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

ARLExisting <- read.csv("Arlington_existing.csv", check.names = FALSE) %>%
  select(Date, `28E`) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

MARExisting <- read.csv("Marshfield_existing.csv", check.names = FALSE) %>%
  select(Date, Marshfield) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

Merged <- list(HANExisting, WMARSExisting, ARLExisting, MARExisting) %>%
  reduce(merge, by = "Date", all = TRUE) %>%
  gather(`S13-S16`, M1400, `28E`, Marshfield, key = "Field", value = "GDDs") %>%
  mutate(Field = case_when(
    Field == "S13-S16" ~ "Hancock (S13-S16)",
    Field == "M1400" ~ "West Madison (M1400)",
    Field == "28E" ~ "Arlington (28E)",
    TRUE ~ Field
  )) %>%
  mutate(Field = factor(Field, levels = c("Hancock (S13-S16)", "West Madison (M1400)", "Arlington (28E)", "Marshfield")))


ggplot(Merged, aes(Date, GDDs, group=Field, color=Field)) +
  geom_line() +
  theme_classic() +
  ylab("Cumulative GDDs (ºF)") +
  ggtitle("Grain Trials") +
  geom_hline(yintercept = 1200, linetype = 'dotted', col = 'blue') +
  annotate("text", x = as.Date("06/10/2023", "%m/%d/%y"), y = 1300, label = "~ start of flowering") +
  geom_hline(yintercept = 120, linetype = 'dotted', col = 'blue') +
  #annotate("text", x = as.Date("06/10/2023", "%m/%d/%y"), y = 170, label = "Approximate time of emergence") +
  scale_color_manual(values = c("#F8766D", "#00BA38", "#619CFF", "black")) +
  geom_hline(yintercept = 2700, linetype = 'dotted', col = 'black') +
  annotate("text", x = as.Date("06/10/2022", "%m/%d/%y"), y = 2550, label = "~ physiological maturity")

ggsave("TrialGDDs.png", dpi=200, width = 6, height = 4, units = "in")
```


### Silage

```{r}

WMARSExisting <- read.csv("WMARS_existing.csv") %>%
  select(Date, M1500) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

ARLExisting <- read.csv("Arlington_existing.csv", check.names = FALSE) %>%
  select(Date, `25N`) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

Merged <- merge(WMARSExisting, ARLExisting, by = "Date") %>%
  gather(M1500, `25N`, key = "Field", value = "GDDs") %>%
  mutate(Field = case_when(
    Field == "M1500" ~ "West Madison (M1500)",
    Field == "25N" ~ "Arlington (25N)"
  ))

Merged$Field <- factor(Merged$Field, levels = c("West Madison (M1500)", "Arlington (25N)"))

ggplot(Merged, aes(Date, GDDs, group=Field, color=Field)) +
  geom_line() +
  theme_classic() +
  ylab("Cumulative GDDs (ºF)") +
  ggtitle("Silage Trials") +
  geom_hline(yintercept = 2195, linetype = 'dotted', col = 'blue') +
  annotate("text", x = as.Date("06/10/2023", "%m/%d/%y"), y = 2050, label = "~ 35% dry matter silage") +
  scale_color_manual(values = c("#00BA38", "#619CFF", "black")) #+
  #scale_x_date(limits = c(as.Date("05/1/2023", "%m/%d/%y"), as.Date("05/25/2023", "%m/%d/%y")))



ggsave("Silage.png", dpi=200, width = 6, height = 4, units = "in")

```

### Nursery

```{r}
WMARSExisting <- read.csv("WMARS_existing.csv") %>%
  select(Date, H703, B501, B601, H702, CERCA) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y")) %>%
  gather(H703, B501, B601, H702, CERCA, key = "Field", value = "GDDs") %>%
  mutate(Field = ifelse(Field == "B501", "B501, B502, H101", Field))

WMARSExisting$Field <- factor(WMARSExisting$Field, levels = c("H703","B501, B502, H101", "B601", "H702", "CERCA"))

ggplot(WMARSExisting, aes(Date, GDDs, group=Field, color=Field)) +
  geom_line() + 
  theme_classic() +
  ylab("Cumulative GDDs (ºF)") +
  ggtitle("Nurseries and Observations") +
  geom_hline(yintercept = 1200, linetype = 'dotted', col = 'blue') +
  annotate("text", x = as.Date("06/10/2023", "%m/%d/%y"), y = 1350, label = "~ start of flowering") +
  geom_hline(yintercept = 120, linetype = 'dotted', col = 'blue') +
  geom_hline(yintercept = 2700, linetype = 'dotted', col = 'black') +
  annotate("text", x = as.Date("06/10/2023", "%m/%d/%y"), y = 2550, label = "~ physiological maturity") +
  scale_color_manual(values = c("#F8766D", "#00BA38", "#619CFF", "black", "gray"))
  #scale_x_date(limits = c(as.Date("05/1/2023", "%m/%d/%y"), as.Date("05/25/2023", "%m/%d/%y"))) #+
  #annotate("text", x = as.Date("06/10/2023", "%m/%d/%y"), y = 170, label = "Approximate time of emergence")

ggsave("NurseriesGDDs.png", dpi=500, width = 6, height = 4, units = "in")
```

### Rainfall

```{r}
HANExisting <- read.csv("Hancock_existing.csv", check.names = FALSE) %>%
  select(Date, Hancock = `Cumulative Rainfall (in)`) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

WMARSExisting <- read.csv("WMARS_existing.csv", check.names = FALSE) %>%
  select(Date, `West Madison` = `Cumulative Rainfall (in)`) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

MARExisting <- read.csv("Marshfield_existing.csv", check.names = FALSE) %>%
  select(Date, Marshfield = `Cumulative Rainfall (in)`) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

ARLExisting <- read.csv("Arlington_existing.csv", check.names = FALSE) %>%
  select(Date, `Arlington` = `Cumulative Rainfall (in)`) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

Merged <- list(HANExisting, WMARSExisting, MARExisting, ARLExisting) %>%
  reduce(merge, by = "Date", all = TRUE) %>%
  gather(Hancock, `West Madison`, Arlington, Marshfield, key = "Location", value = "Rainfall")

Merged$Location <- factor(Merged$Location, levels = c("Hancock", "West Madison", "Arlington", "Marshfield"))

ggplot(Merged, aes(Date, Rainfall, group=Location, color=Location)) +
  geom_line() +
  theme_classic() +
  ylab("Cumulative rainfall (in.)") +
  scale_color_manual(values = c("#F8766D", "#00BA38", "#619CFF", "black")) +
  #scale_x_date(limits = c(as.Date("04/25/2023", "%m/%d/%y"), as.Date("05/25/2023", "%m/%d/%y"))) +
  #scale_y_continuous(limits = c(0, 5)) +
  labs(title = "Rainfall", subtitle = "Note: WM rainfall periods start and end at 6 am rather than midnight")

ggsave("Rainfall.png", dpi=200, width = 6, height = 4, units = "in")
```





Here's the code to run this from the command line
```
Rscript -e "rmarkdown::render('GDD-Plots-2023.Rmd')"
```